Left-linear rules in adhesive categories

### 2024-02-13 icalp/fusione.tex

@@

\todo[inline]{Forse non introdurrei neppure la classe dei strong enforcing, tanto usiamo solo i well switching, o no?}

\todo[inline]{Io invece la introdurrei proprio come punto teorico che mi sembra parecchio importante: qua li condizioni mi garantiscono che gli independence pair implicano la scambiabilità? [DC]}

@@

\begin{restatable}[confluence of switchings]{lemma}{lemSwitchConfluence}
  \label{lem:switch-confluence}
  Let $\X$ be an $\mathcal{M}$-adhesive category and $(\X,\R)$ a
  left-linear DPO rewriting system.
  % 
  Consider a derivation $\der{D}=\{\dder{D}_i\}_{i=0}^2$ and suppose
  that $(i_0,i_1)$ is a strong independence pair between $\dder{D}_0$
  and $\dder{D}_1$, $(a_0,a_1)$ one between $\dder{D}_1$ and
  $\dder{D}_2$ and $(e_0, e_1)$ one between $\dder{D}_0$ and
  $S_{a_0,a_1}(\dder{D}_2)$.\todo{P: non e' meglio qui limitarci a well switching, cosi' da evitare riferimenti a strong pairs?}
\end{restatable}

@@

\begin{restatable}[no need of useless switchings]{theorem}{thNoNeed}
  \label{th:no-need}
  Let $\der{D}$ and $\der{D}'$ be derivation sequences. If $\der{D} \shifteq \der{D}'$ then there is a switching sequence consisting only of inversions
  \begin{center}
    $\der{D} = \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n=\der{D}'$
  \end{center}
\end{restatable}

@@

\begin{restatable}[canonical form]{corollary}{coCanonical}
  \label{co:canonical}
  Let $\der{D}$ and $\der{D}'$ be derivation sequences. If
  $\der{D} \shifteq[\sigma]\der{D}'$ then there is a switching sequence
  \begin{center}
    $\der{D} = \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n=\der{D}'$
  \end{center}
  where for each $i \in \interval{n}$ the transpostion $\nu_i  = \transp{k}$ is such that 
  $k = \max \{ j \mid \transp{j} \in \inv{\nu_{i,n}}\}$.\todo{P: $\nu_{i,n} = \sigma \circ \nu_{1,i}$, giusto? forse e' piu' carino}
\end{restatable}

>>>>

\begin{notation} We will denote any switch of two switchable direct derivations $\dder{D}_0$ and $\dder{D}_1$ along $(i_0, i_1)$ by $S_{i_0, i_1}(\dder{D}_0, \dder{D}_1)$. Moreover, we will use $S_{i_0,i_1}(\dder{D}_1)$ and  $S_{i_0,i_1}(\dder{D}_0)$ to denote its components, in such a way that \[S_{i_0, i_1}(\dder{D}_0, \dder{D}_1)=S_{i_0,i_1}(\dder{D}_1)\cdot S_{i_0,i_1}(\dder{D}_0)\]
\end{notation}

@@

@@

\begin{restatable}[Confluence of switchings]{lemma}{lemSwitchConfluence}
  \label{lem:switch-confluence}
  Let $\X$ be an $\mathcal{M}$-adhesive category and $(\X,\R)$ a
  left-linear DPO rewriting system.
  % 
  Consider a derivation $\der{D}=\{\dder{D}_i\}_{i=0}^2$ and suppose
  that $(i_0,i_1)$ is a strong independence pair between $\dder{D}_0$
  and $\dder{D}_1$, $(a_0,a_1)$ one between $\dder{D}_1$ and
  $\dder{D}_2$ and $(e_0, e_1)$ one between $\dder{D}_0$ and
  $S_{a_0,a_1}(\dder{D}_2)$.\todo{P: non e' meglio qui limitarci a well switching, cosi' da evitare riferimenti a strong pairs?}
\end{restatable}

@@

\begin{restatable}[No need of useless switchings]{theorem}{thNoNeed}
  \label{th:no-need}
  Let $\der{D}$ and $\der{D}'$ be derivation sequences. If $\der{D} \shifteq \der{D}'$ then there is a switching sequence consisting only of inversions
  \begin{center}
    $\der{D} = \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n=\der{D}'$
  \end{center}
\end{restatable}

@@

\begin{restatable}[Canonical form]{corollary}{coCanonical}
  \label{co:canonical}
  Let $\der{D}$ and $\der{D}'$ be derivation sequences. If
  $\der{D} \shifteq[\sigma]\der{D}'$ then there is a switching sequence
  \begin{center}
    $\der{D} = \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n=\der{D}'$
  \end{center}
  where for each $i \in \interval{n}$ the transpostion $\nu_i  = \transp{k}$ is such that 
  $k = \max \{ j \mid \transp{j} \in \inv{\nu_{i,n}}\}$.\todo{P: $\nu_{i,n} = \sigma \circ \nu_{1,i}$, giusto? forse e' piu' carino}
\end{restatable}
